Generalization Errors of Hard Parameter Sharing: High-dimensional Asymptotics and Implications

### 2020-09-15 arxiv/diff_cov.tex
sec b.2
wip

\begin{definition}[Limiting bias]
Recall that $a_1$ and $a_2$ are the solutions of the equation system \eqref{eq_a12extra}. Let
	\[ \Pi_\bias \define  M \frac{a_3 M^{\top}M + (1 + a_4) \id_{p\times p}}{(a_1 M^{\top}M + a_2\id_{p\times p})^2} M^{\top}, \]
where $a_{3}$ and $a_4$ are the solutions of the following two equations:
\begin{gather}\label{eq_a34extra}
		\left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
		= b_0, \quad \left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1 .
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
	\end{gather}
In the above equation, we denote $b_k$ as $\frac1{p}\sum_{i=1}^p {\lambda_i^{2k}} / { (\lambda_i^2 a_1 + a_2)^2  }$, for $k = 0, 1, 2$.
\end{definition}

@@

Lemma \ref{lem_cov_shift} also implies the asymptotic limit of the prediction loss of single-task learning.
%The next lemma, which is , helps to determine the asymptotic limit of $\te(\hat{\beta}_t^{\STL})=\sigma^2   \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}$ as $p\to \infty$.
In particular, by setting $X_1 = 0$, we obtain the following corollary, which is a well-known result in random matrix theory.
\begin{corollary}[See e.g. Theorem 2.4 in \citet{isotropic}]\label{lem_minv}
	In the setting of Lemma \ref{lem_cov_shift}, with high probability we have that
	\be\label{XXA}  \bigtr{(X_2^{\top}X_2)^{-1}\Sigma} = \frac{1}{n_2 - p} \cdot \bigtr{\Sigma_2^{-1}\Sigma} + o(\norm{\Sigma}). \ee
\end{corollary} %\bigo{ \|A\|p^{-1/2+\epsilon}}
To see this, note that when $n_1=0$, $a_1 = 0$ and $a_2 = (n_2-p) / n_2$ is the solution of \eqref{eq_a12extra}, and one can see that equation \eqref{lem_cov_shift_eq} reduces to equation \eqref{XXA}.
We briefly describe the history behind the above well-known result.
When the entries of $X_2$ are multivariate Gaussian, this result recovers the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For general non-Gaussian random matrices, it can be obtained using Stieltjes transform method; see e.g., Lemma 3.11 of \cite{bai2009spectral}.
Here we have stated a result from Theorem 2.4 in \cite{isotropic}, which gives an almost sharp concentration error bound.
One can see that our result extends Lemma \ref{lem_minv} from a single sample covariance matrix to the sum of two independent sample covariance matrices.

>>>>

\begin{definition}[Limiting bias]
Recall that $a_1$ and $a_2$ are the solutions of the equation system \eqref{eq_a12extra}. Let
	\[ \Pi_\bias \define  M \frac{a_3 M^{\top}M + (1 + a_4) \id_{p\times p}}{(a_1 M^{\top}M + a_2\id_{p\times p})^2} M^{\top}, \]
where $a_{3}$ and $a_4$ are the solutions of the following two equations:
\begin{gather}\label{eq_a34extra}
		\left(\frac{\rho_1}{a_1^{2}} -  b_2  \right)\cdot  a_3 -  b_1 \cdot  a_4 = b_1,\quad \left(\frac{\rho_2}{a_2^{2}}-  b_0\right)\cdot  a_4 - b_1 \cdot  a_3
		= b_0 .
%		\left(\frac{n_1}{\hat a_1^2} -  \sum_{i=1}^p \frac{\hat \lambda_i^4   }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_4 -\left(\sum_{i=1}^p \frac{\hat \lambda_i^2  }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }\right)\hat a_3
%		= \sum_{i=1}^p \frac{\hat \lambda_i^2 }{  (\hat a_2 + \hat \lambda_i^2\hat a_1)^2  }. \label{eq_a4}
	\end{gather}
In the above equation, we denote $b_k$ as $\frac1{p}\sum_{i=1}^p {\lambda_i^{2k}} / { (\lambda_i^2 a_1 + a_2)^2  }$, for $k = 0, 1, 2$.
\end{definition}

@@

Lemma \ref{lem_cov_shift} also implies the asymptotic limit of the prediction loss of single-task learning.
%The next lemma, which is , helps to determine the asymptotic limit of $\te(\hat{\beta}_t^{\STL})=\sigma^2   \bigtr{(X_2^{\top}X_2)^{-1}\Sigma_2}$ as $p\to \infty$.
In particular, by setting $X_1 = 0$, we obtain Lemma \ref{lem_minv}.
%\begin{corollary}[See e.g. Theorem 2.4 in \citet{isotropic}]\label{lem_minv}
%	In the setting of Lemma \ref{lem_cov_shift}, with high probability we have that
%	\be\label{XXA}  \bigtr{(X_2^{\top}X_2)^{-1}\Sigma} = \frac{1}{n_2 - p} \cdot \bigtr{\Sigma_2^{-1}\Sigma} + o(\norm{\Sigma}). \ee
%\end{corollary} %\bigo{ \|A\|p^{-1/2+\epsilon}}
To see this, note that when $n_1=0$, $a_1 = 0$ and $a_2 = (n_2-p) / n_2$ is the solution of \eqref{eq_a12extra}, and one can see that equation \eqref{lem_cov_shift_eq} reduces to equation \eqref{XXA}.
We briefly describe the history behind the above well-known result.
When the entries of $X_2$ are multivariate Gaussian, this result recovers the classical result for the mean of inverse Wishart distribution \cite{anderson1958introduction}.
For general non-Gaussian random matrices, it can be obtained using Stieltjes transform method; see e.g., Lemma 3.11 of \cite{bai2009spectral}.
Here we have stated a result from Theorem 2.4 in \cite{isotropic}, which gives an almost sharp concentration error bound.
One can see that our result extends Lemma \ref{lem_minv} from a single sample covariance matrix to the sum of two independent sample covariance matrices.

### 2020-09-15 arxiv/setup.tex
sec b.2
wip

\subsection{Random Matrices}
\todo{merge w/ theorem}
We make several standard assumptions on the covariance matrix $\Sigma_k$ and the random vector $z_i$, for $k = 1,\dots, t$ and $i = 1,\dots, n_k$.
First, for every entry of $z_i$, we assume that its $(4+\e)$-th moment exists, that is, there exist constants $\e, C > 0$ such that
\begin{equation}\label{assmAhigh}
	\ex{\abs{z_{i,j}}^{4+\e}} \le C,\ 1\le j \le p.
\end{equation}
Second, we assume that the spectral norm of $\Sigma_k$ is bounded by a fixed constant, for every $k = 1,\dots, t$.
This is without loss of generality since we can always rescale $\Sigma_k$ so that its spectral norm satisfies the condition.
Finally, for the noise vector $\varepsilon_k$, we assume that there exists an absolute constant such that for all $l$ \todo{check, remove $c$ and $o(1)$ in main body}, the $l$-th moment of the noise entries exists:
\begin{align*}
	\ex{\abs{\varepsilon_{k, j}}^l \le C}, \text{ for all } j = 1, \dots, n_k.
\end{align*}

>>>>

\subsection{Random Matrices}

We introduce several standard assumptions on the random matrices we study.
\begin{assumption}\label{assume_rm}
	Let $c > 4$ be a fixed value.
	Let $X = Z \Sigma^{1/2} \in\real^{n\times p}$ be a random matrix where $Z \in \real^{n\times p}$ consists of i.i.d. entries with zero mean and unit variance and $\Sigma \in \real^{p\times p}$ is a positive semidefinite matrix.
	We assume that $\Sigma$ and $Z$ satisfy the following (standard) conditions, respectively.
	\begin{enumerate}
		\item[i)] Bounded moments: for every entry of $Z$, we assume that its $c$-th moment exists, that is, there exist a fixed constant $C > 0$ such that
			\begin{equation}\label{assmAhigh}
				\ex{\abs{Z_{i,j}}^{c}} \le C, \text{ for any } 1\le i \le n, 1\le j \le p.
			\end{equation}
		\item[ii)] Bounded singular values: the singular values of $\Sigma$ are bounded within a fixed range, that is, there exists two fixed values $C > C' > 0$ such that all of $\Sigma$'s singular values are bounded within $C'$ and $C$.
	\end{enumerate}
\end{assumption}

\begin{lemma}[cf. Theorem 2.4 in \citet{isotropic}]\label{lem_minv}
	Let $X  \in \real^{n\times p}$ be a random matrix that satisfies Assumption \ref{assume_rm}.
	Let $\Sigma\in\real^{p\times p}$ denote the population covariance matrix of $X$.
	Let $\e$ be any fixed value within $(0, \frac{c - 4}{2c})$.
	With high probability over the randomness of $X$, we have that
	\[ \bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{\sigma^2 \cdot p}{n - p} + \OO(p^{-\e}). \]
\end{lemma}

%Second, we assume that the spectral norm of $\Sigma_k$ is bounded by a fixed constant, for every $k = 1,\dots, t$.
%This is without loss of generality since we can always rescale $\Sigma_k$ so that its spectral norm satisfies the condition.
Finally, for the random noise of every task, we assume that all $l$, the $l$-th moment of the noise entries exists:
\begin{align}\label{assmAhigh2}
	\ex{\abs{\varepsilon_{k, j}}^l \le C}, \text{ for all } j = 1, \dots, n_k.
\end{align}
